---
title: 'Datacamp: Intermediate R'
output:
  word_document: default
  pdf_document: default
  html_notebook: default
---
Note:
• Learn more about R notebooks here: https://support.posit.co/hc/en-us/articles/200552276-Creating-Notebooks-from-R-Scripts-in-the-RStudio-IDE
• This is a dump notebook of all the useful stuff I learned seperated by modules.

#
# Intermediate R Notes
#
Below are selected notes and examples from the DataCamp Intermediate R course.

# Functions - Intermediate R
```{r}
values <- c(1,4,7, 9, NA, 10, 19)
sd(values)
# notice that prints NA because of the NA in vector
sd(x = values, na.rm = TRUE)
# By contrast, if na.rm - FALSE...
sd(x = values, na.rm = FALSE) #...we get NA again
```
You can also trim the values and calculate the mean (or use an function) to get the mean of muiltiple variables, as well as trim the result:

```{r}
# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# Calculate the mean of the sum
avg_sum <- mean(linkedin + facebook)

# Calculate the trimmed mean of the sum
avg_sum_trimmed <- mean(linkedin + facebook, trim = 0.2, na.rm = FALSE)

# Inspect both new variables
print(avg_sum)
print(avg_sum_trimmed)
```

Another example of a function is the `apply()` function. It allows you to apply a function to the rows or columns of a 
matrix or data frame. The syntax is as follows:
 
# Syntax:
# apply(X, MARGIN, FUN, ...)
Where:
- `X` is the matrix or data frame
- `MARGIN` is either 1 (for rows) or 2 (for columns)
- `FUN` is the function to apply
- `...` are any additional arguments to pass to the function


# IMPORTANT: R passes arguments by value
# The title gives it away already: R passes arguments by value. What does this mean? 
# Simply put, it means that an R function cannot change the variable that you input to that function.

# Using Packaages: https://campus.datacamp.com/courses/intermediate-r/chapter-3-functions?ex=16
Use the search() function to find the function you want to use.
```{r}
#instal the package
install.packages("ggplot2")

# Load the ggplot2 package
library(ggplot2)

# Retry the qplot() function
qplot(mtcars$wt, mtcars$hp)

# Check out the currently attached packages again
search()
```
• R loads 7 packages by default: base, utils, datasets, grDevices, graphics, stats, and methods.
• require() = loads packages into your R session

There are muiltiple ways to load packages:
```{r}
# Chunk 1
library(data.table)
require(rjson)

# Chunk 2
library("data.table")
require(rjson)
```
These two options list all of the chunks that do not generate an error.

# Useful math functions that R features:
• abs(): Calculate the absolute value.
• sum(): Calculate the sum of all the values in a data structure.
• mean(): Calculate the arithmetic mean.
• round(): Round the values to 0 decimal places by default. 

# Data Utilities
R features a bunch of functions to juggle around with data structures::
• seq(): Generate sequences, by specifying the from, to, and by arguments.
• rep(): Replicate elements of vectors and lists.
• sort(): Sort a vector in ascending order. Works on numerics, but also on character strings and logicals.
• rev(): Reverse the elements in a data structures for which reversal is defined.
• str(): Display the structure of any R object.
• append(): Merge vectors or lists.
• is.*(): Check for the class of an R object.
• as.*(): Convert an R object from one class to another.
• list(): Flatten (possibly embedded) lists to produce a vector.

# Regular expressions
Regular expressions (regex) are a powerful tool for pattern matching 
and manipulation of strings in R. They allow you to search for specific 
patterns within text, replace text, or split text based on patterns.

• grepl(): The primary function for working with regular expressions in R.
• grep(): Searches for matches to a pattern within a character vector and returns the 
          indices of the matches.

# grepl & grep (2) - DATA CLEANING
You can use the caret, ^, and the dollar sign, $ to match the content 
located in the start and end of a string, respectively. This could take 
us one step closer to a correct pattern for matching only the ".edu" email 
addresses from our list of emails. But there's more that can be added to make the 
pattern more robust:

• @, because a valid email must contain an at-sign.
• .*, which matches any character (.) zero or more times (*). Both the dot 
  and the asterisk are metacharacters. You can use them to match any character 
  between the at-sign and the ".edu" portion of an email address.
• \\.edu$, to match the ".edu" part of the email at the end of the string. 
  The \\ part escapes the dot: it tells R that you want to use the . as an actual 
    character.

```{r}
# sub() and gsub() are used for replacing patterns in strings

# sub() replaces the first match of the pattern
# sub(pattern, replacement, x, ignore.case = FALSE, perl = FALSE, fixed = FALSE, useBytes = FALSE)

# gsub() replaces all matches of the pattern
# gsub(pattern, replacement, x, ignore.case = FALSE, perl = FALSE, fixed = FALSE, useBytes = FALSE)

# Example of using sub() and gsub() to replace email domains ewith ".datacamp.edu":
emails <- c("jdoe@wisc.edu", "hdong@msu.edu", "tchapman@iu.edu")
sub(pattern = ".*@.*\\.edu$", replacement = "datacamp.edu", x = emails)
```

# Times & Dates 
R has a built-in class for dates and times called POSIXct.
• POSIXct: Represents the number of seconds since the epoch (1970-01-01 00:00:00 UTC).
• POSIXlt: Represents a list of components (year, month, day, hour, minute, second) 
  for a specific date and time.
• Sys.Date(): Returns the current date in the POSIXct format.
• Sys.time(): Returns the current date and time in the POSIXct format.
• as.Date(): Converts a character string to a date object.
• as.POSIXct(): Converts a character string to a POSIXct object.  

Let's try a little exercise to calculate the difference between two dates (first date anniversary--shoutout pookie bear)!
```{r}
anniversary <- as.Date("2022-11-18")
#get the current date
today <- Sys.Date()
# Calculate the difference in days
diff <- today - anniversary
# Print the difference
print(cat("It has been ", diff, " days since the anniversary."))
# Now how about in years?
diff_years <- as.numeric(diff) / 365
# Print the difference in years
print(cat("It has been ", diff_years, " years since the anniversary."))
```
Note that under the hood, R is able to do the date arithmetic easily because
it converts the dates into a number based on the difference between the given 
date and the epoch date (1970-01-01 at midnight or 00:00:00).
```{r}
my_date <- as.Date("2022-11-18")
unclass(my_date)
# You should see the following output in the console:
# [1] 19263
# The number of days since the epoch date (Jan 1, 1970)
```

# Create and format dates
To create a Date object from a simple character string in R, you can use the as.Date() function. The character string has to obey a format that can be defined using a set of symbols (the examples correspond to 13 January, 1982):
----------------THIS IS REALLY IMPORTANT!!!!----------------
                %Y: 4-digit year (1982)
                %y: 2-digit year (82)
                %m: 2-digit month (01)
                %d: 2-digit day of the month (13)
                %A: weekday (Wednesday)
                %a: abbreviated weekday (Wed)
                %B: month (January)
                %b: abbreviated month (Jan)
----------------THIS IS REALLY IMPORTANT!!!!----------------

The following R commands will all create the same Date object for the 13th day in 
January of 1982:
```{r}
as.Date("1982-01-13")
as.Date("Jan-13-82", format = "%b-%d-%y")
as.Date("13 January, 1982", format = "%d %B, %Y")
```
Notice that the first line here did not need a format argument, because by default 
R matches your character string to the formats "%Y-%m-%d" or "%Y/%m/%d".

In addition to creating dates, you can also convert dates to character strings that 
use a different date notation. For this, you use the format() function. Try the following 
lines of code:
```{r}
today <- Sys.Date()
format(Sys.Date(), format = "%d %B, %Y")
format(Sys.Date(), format = "Today is a %A!")
```

Here is a funky exercise that datacamp threw at me (instrctions below):

• Three character strings representing dates have been created for you. Convert them 
to dates using as.Date(), and assign them to date1, date2, and date3 respectively. 
The code for date1 is already included.
• Extract useful information from the dates as character strings using format(). 
From the first date, select the weekday. From the second date, select the day of the 
month. From the third date, you should select the abbreviated month and the 4-digit year, 
separated by a space.
```{r}
# Definition of character strings representing dates
str1 <- "May 23, '96"
str2 <- "2012-03-15"
str3 <- "30/January/2006"

# Convert the strings to dates: date1, date2, date3
date1 <- as.Date(str1, format = "%b %d, '%y")
date2 <- as.Date(str2, format = "%Y-%m-%d")
date3 <- as.Date(str3, format = "%d/%B/%Y")


# Convert dates to formatted strings
format(date1, "%A")
format(date2, "%d")
format(date3, "%b %Y")
```
# Create and format times
Similar to working with dates, you can use as.POSIXct() to convert from a 
character string to a POSIXct object, and format() to convert from a POSIXct 
object to a character string. Again, you have a wide variety of symbols:
----------------THIS IS REALLY IMPORTANT!!!!----------------
              %H: hours as a decimal number (00-23)
              %I: hours as a decimal number (01-12)
              %M: minutes as a decimal number
              %S: seconds as a decimal number
              %T: shorthand notation for the typical format %H:%M:%S
              %p: AM/PM indicator
----------------THIS IS REALLY IMPORTANT!!!!----------------
For a full list of conversion symbols, consult the strptime documentation
