---
title: 'Datacamp: Intermediate R'
output:
  word_document: default
  pdf_document: default
  html_notebook: default
---
Note:
• Learn more about R notebooks here: https://support.posit.co/hc/en-us/articles/200552276-Creating-Notebooks-from-R-Scripts-in-the-RStudio-IDE
• This is a dump notebook of all the useful stuff I learned seperated by modules.

#
# Intermediate R Notes
#
Below are selected notes and examples from the DataCamp Intermediate R course.

# Functions - Intermediate R
```{r}
values <- c(1,4,7, 9, NA, 10, 19)
sd(values)
# notice that prints NA because of the NA in vector
sd(x = values, na.rm = TRUE)
# By contrast, if na.rm - FALSE...
sd(x = values, na.rm = FALSE) #...we get NA again
```
You can also trim the values and calculate the mean (or use an function) to get the mean of muiltiple variables, as well as trim the result:

```{r}
# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# Calculate the mean of the sum
avg_sum <- mean(linkedin + facebook)

# Calculate the trimmed mean of the sum
avg_sum_trimmed <- mean(linkedin + facebook, trim = 0.2, na.rm = FALSE)

# Inspect both new variables
print(avg_sum)
print(avg_sum_trimmed)
```

Another example of a function is the `apply()` function. It allows you to apply a function to the rows or columns of a 
matrix or data frame. The syntax is as follows:
 
# Syntax:
# apply(X, MARGIN, FUN, ...)
Where:
- `X` is the matrix or data frame
- `MARGIN` is either 1 (for rows) or 2 (for columns)
- `FUN` is the function to apply
- `...` are any additional arguments to pass to the function


# IMPORTANT: R passes arguments by value
# The title gives it away already: R passes arguments by value. What does this mean? 
# Simply put, it means that an R function cannot change the variable that you input to that function.

# Using Packaages: https://campus.datacamp.com/courses/intermediate-r/chapter-3-functions?ex=16
Use the search() function to find the function you want to use.
```{r}
#instal the package
install.packages("ggplot2")

# Load the ggplot2 package
library(ggplot2)

# Retry the qplot() function
qplot(mtcars$wt, mtcars$hp)

# Check out the currently attached packages again
search()
```
• R loads 7 packages by default: base, utils, datasets, grDevices, graphics, stats, and methods.
• require() = loads packages into your R session

There are muiltiple ways to load packages:
```{r}
# Chunk 1
library(data.table)
require(rjson)

# Chunk 2
library("data.table")
require(rjson)
```
These two options list all of the chunks that do not generate an error.

# Useful math functions that R features:
• abs(): Calculate the absolute value.
• sum(): Calculate the sum of all the values in a data structure.
• mean(): Calculate the arithmetic mean.
• round(): Round the values to 0 decimal places by default. 